---
title: "Shiny Project - IMDB - Data Preparation Step 1"
author: "Mitch Abramson"
date: "July, 2017"
output: html_document
---

### Notes

- IMDB has a list of the top [250 movies](http://www.imdb.com/chart/top). 
- This file contains code extracted and updated from a previous effort for use in the shiny_IMDB project.
- As per comments, multiple "hot list" on IMDB, (top 250, bottom 100, ...) were extracted from the web using this code

```{r sourceDataProb1, warning=F, message=F}

library(RCurl)  # RCurl might be needed for this code
library(XML)    # XML is the main workhorse of parsing HTML pages through XPath to extract specific fields
                # inputs are assumed to be HTML pages with tables on them
library(dplyr)  # need dplyer or magrittr for %>%

# uncomment a URL and uncomment other URLs to scrape a particular data set
# last URL used with this script is uncommented

# movies_url       <- "http://www.imdb.com/chart/top"  # renamed to head off conflict w/ base 'url'
# movies_url       <- "http://www.imdb.com/chart/bottom" 
# movies_url       <- "http://www.imdb.com/chart/top-english-movies"
movies_url         <- "http://www.imdb.com/india/top-rated-indian-movies/"

```
Get the data.  Need help with the xpath?  Here are the links I used:

  - [w3c exaplains substring functions](https://www.w3.org/TR/1999/REC-xpath-19991116/#function-substring-before)
  - [stackoverflow example of substring-before / substring-after](http://stackoverflow.com/questions/4720480/xpath-select-only-part-of-an-attribute)

Analysis of the target HTML:

  - movie name: \<td class="titleColumn"\>\<a\>Movie Title ...
  - movie release yr: \<td class="titleColumn"\>\<span class="secondaryInfo"\>(1994)\</span\>
  - movie ranking: \<td class="ratingColumn imdbRating"\>\<strong title="meta info"\>9.2\</strong\>
  - movie ranking weight: ...\<strong title="9.2 based on 1,771,990 user ratings"\>...
  - movie id: <td class="ratingColumn"><div ... data-titleid="tt0111161"> (first div)

```{r, echo=T, eval=T}
# required libraries loaded in previous cell

print("Inputs to Data Extraction:")
movies_url    # show url being used in code that follows

IMDB_mvNm_xPath <- '//td[@class="titleColumn"]/a'
IMDB_mvYr_xPath <- '//td[@class="titleColumn"]/span[@class="secondaryInfo"]'
IMDB_mvRk_xPath <- '//td[@class="ratingColumn imdbRating"]/strong'
IMDB_mvID_xPath <- '//td[@class="ratingColumn"]/div[1]'

# show xpaths ahead of using them:
IMDB_mvNm_xPath
IMDB_mvYr_xPath
IMDB_mvRk_xPath
IMDB_mvID_xPath

```

```{r}

# function to create each cell of our data set from the IMDB HTML:

vectorize_xml_webCell <- function(src_url, src_xpath) {
  htmlDataVector = c()
  htmlDataVector <- getURL(src_url) %>%
                    htmlParse(.) %>%
                    getNodeSet(., src_xpath) %>%
                    sapply(., xmlValue)
  
  return(htmlDataVector)
}

```

```{r}

vectorize_xml_webCellAttr <- function(src_url, src_xpath, attrName) {
  htmlDataVector = c()
  htmlDataVector <- getURL(src_url) %>%
                    htmlParse(.) %>%
                    getNodeSet(., src_xpath) %>%
                    sapply(., xmlGetAttr, attrName)
                    # xmlGetAttr is the function to extract data from the attrName on the node set
  
  return(htmlDataVector)
}

```

```{r movieVectors, echo=T, eval=T}

movie_names      <- vectorize_xml_webCell(movies_url, IMDB_mvNm_xPath)
movie_release_yr <- vectorize_xml_webCell(movies_url, IMDB_mvYr_xPath) %>%
                    substring(., 2, nchar(.)-1) %>%
                    as.integer(.)
movie_rating     <- vectorize_xml_webCell(movies_url, IMDB_mvRk_xPath) %>%
                    as.double(.)                                     # convert to numbers
movie_rating_wt  <- vectorize_xml_webCellAttr(movies_url, IMDB_mvRk_xPath, "title") %>%
                    gsub("\\d.\\d based on\\s*", "", .) %>%
                    gsub("\\s*user ratings\\s*", "", .) %>%
                    gsub("\\,", "", .) %>%
                    as.integer(.)
                    
             # how many ratings went into the result
             # extra info Mitch would want if doing this project in the real world

movie_ID         <- vectorize_xml_webCellAttr(movies_url, IMDB_mvID_xPath, "data-titleid") 
             # this should help with problem 5 and beyond

# test the results ... let's just look at 5 return values and the count from each item
print("Quick test of the data:")
movie_names[1:5]
movie_release_yr[1:5]
movie_rating[1:5]
movie_rating_wt[1:5]
movie_ID[1:5]

```

```{r}

# think of problem1 as the data scraping problem to solve
# problem1_data is the output of this script to write to csv for use in data projects

problem1_data = data.frame(
  Movie_Title = movie_names,
  YR_Released = movie_release_yr,
  Rating = movie_rating,
  Num_Reviews = movie_rating_wt,
  Movie_ID = movie_ID,
  stringsAsFactors = FALSE
)

head(problem1_data)

```

```{r wholeDF, echo=FALSE, eval=FALSE}
problem1_data

```

```{r}

# if not using hard coded paths - seeing current working directory before performing write is a good idea
# fullFilePath variable below may negate need for this cell

getwd()

```

```{r}
# save data frame to csv:

# set shiny project path on computer here:
# ------------------------------------------
fileDir <- "/Users/mitchmac/Documents/DS_BootCamp/projects/shiny_imdb/data/"

# 
# fullFilePath <- paste0(fileDir, 'IMDB_Top250movies.csv')
# fullFilePath <- paste0(fileDir, 'IMDB_Bottom250movies.csv')
# fullFilePath <- paste0(fileDir, 'IMDB_Top250Engmovies.csv')
fullFilePath <- paste0(fileDir, 'IMDB_Top250Indianmovies.csv')

write.csv(problem1_data, file=fullFilePath)              

```

Output files shown above were used as inputs to next step in data preparation process.  These files were manually edited
into files of the same name but ending in "2".  The manual edit added a simple column identifying which top or bottom list
data originated from so we can track this later in our process and for use in the project.

A future edit of this code should probably automate this step.

